---
title: "VLH_turbine_090226"
output: html_document
date: "2026-02-09"
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

*Go through this analysis and add descriptions, results etc where
necessary. The document should be able to run visa 'run \> run
all'.....*

**Transfer from processing stage to analysis stage**

Data used throughout this analysis has been processed using the
semi-automated analysis script. 'R_BDS_import.R' The script is designed
to assist the user in processing multiple RAPID data files to extract
key variables on pressure and acceleration.

**Identification & Basic Info**

-   `long_id`: Unique sensor deployment identifier

-   `treatment`:

-   `time`:

-   `pres`:

-   `accmag`:

-   `mean_strain`:

-   `roi`:

-   `configuration`:

*Detail about any process that creates and shapes the data before this
analysis....*

This markdown file is formatted to be executed with a single run by
importing pre-processed BDS sensor data and performing downstream
analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  fig.align = "center"
)
knitr::opts_knit$set(root.dir = getwd())
```

# 1. Required libraries and helpers

```{r libraries, echo=FALSE, message = FALSE}

library(tidyverse) # mainly for dplyr and ggplot
library(plotly) # interactive plot
library(ggdist) # kernel density plot
library(cowplot) #arrange multipanel plots
library(kSamples) # Anderson Darling test
library(binom) #95% CI Wilson
library(dlookr)

```

## 1.1 Import sensor data

*All processed data files corresponding to each treatment condition are
stored in....*

Import and clean pre-processed sensor data. Import data already has 'bed
sensors' removed. Some final smaller changes made to data. See comments
where necessary.

```{r data_import, echo=FALSE, message = FALSE}

denver_files <- list.files(
  "./input",
  pattern = "^denver_.*\\.csv$",
  full.names = TRUE
)

denver_data <- denver_files %>%
  map_dfr(read_csv)


# Setup data structure with expanded treatments
denver_data <- denver_data %>%
  mutate(
    # Extract RPM as numeric for dose-response modeling
    rpm = as.numeric(str_extract(treatment, "\\d+")),
    
    # Extract configuration type
    configuration = factor(
      str_extract(treatment, "(original|modified)"),
      levels = c("original", "modified"),
      labels = c("Original", "Modified")
    ),
    
    treatment = factor(
      treatment,
      levels = c("30rpm_original", "40rpm_original",
                 "50rpm_original", "60rpm_original",
                 "40rpm_modified", "50rpm_modified"),
      labels = c("30 RPM Original", "40 RPM Original", 
                 "50 RPM Original", "60 RPM Original",
                 "40 RPM Modified", "50 RPM Modified")
    ),
    roi = factor(
      roi,
      levels = c("unlabeled", "nadir", "1s_post"),
      labels = c("Unlabelled", "Nadir", "Post-nadir")
    )
  )

# modify 5 sensors... 
denver_data <- denver_data %>%
  filter(
    # Remove C550327103129 entirely
    long_id != "C550327103129",
    # Trim specific sensors
    !(long_id == "U120327084613" & time > 113.51),
    !(long_id == "U100327103204" & time > 422),
    !(long_id == "U160510111356" & time > 85),
    !(long_id == "U190327103259" & time > 290)
  )

```

Input helper functions

```{r helper_functions, echo=FALSE, message = FALSE}

# setup custom theme
theme_custom <- function(base_size=10){ 
  #base font size set here (10 pt)
  theme_grey() %+replace%
    theme(
      axis.text = element_text(colour="black"),
      axis.title = element_text(colour="black"),
      axis.ticks = element_line(colour="black"),
      panel.border = element_rect(colour = "black", fill=NA),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      strip.background = element_rect(colour = "black", fill = NA),
      panel.spacing.x = unit(12, "pt")
      #default family = sans (Arial)
    ) 
}

```

# 2. Exploratory data analysis

All EDA should be provided here. This includes measuring and validating
data dimensions e.g., number sensors for each operational scenario,
cleaning data including missing and outlier data and project specific
conditions, setting definitions for what data is used for downstream
analysis, early summarisations and visualizations which lead into later
formal analysis

E.g.,

```{r data_structure, echo=FALSE, message = FALSE}

#Check unique sensor ID for each treatment
denver_data %>%
  group_by(treatment) %>%
  summarise(
    unique_long_ids = n_distinct(long_id),
    .groups = "drop"
  )

str(denver_data)

colnames(denver_data)

dlookr::plot_outlier(denver_data)

```

## 2.1 Visualize main relationships

By plotting the main visual relationships we're able to quickly identify
relationships in the data and begin addressing study objectives and
hypothesis.

Think of your study objectives.... what do those relationships look
like?

E.g.,

```{r accmag_post_nadir}

plot_data_post <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time),
         mean_strain = round(mean_strain)
         ) %>%
  ungroup()

# Get max acceleration points for each sensor
max_points <- plot_data_post %>%
  group_by(long_id, treatment) %>%
  slice_max(order_by = abs(accmag), n = 1, with_ties = FALSE) %>%
  ungroup()

# accmag_time_plot
accmag_post_nadir <- ggplot(plot_data_post, aes(x = time_post, y = abs(accmag), group = long_id, color = treatment)) +
  geom_line(alpha = 0.5, color = "black") +
  geom_point(data = max_points, alpha = 0.7, size = 2, color = "red2") +
  facet_wrap(~ treatment, nrow = 1, labeller = labeller(treatment = c(
    "Original" = "Original Configuration",
    "Modified" = "Modified Configuration"
  ))) +
  labs(
    x = " Time (s) (ROI 3 Post-nadir)", 
    y = expression("Acceleration Magnitude (m s" ^-2*")")) +
  scale_x_continuous(breaks = seq(0, 1, 0.1), expand = c(0,0))+
  scale_y_continuous(breaks = seq(0, 250, 25), expand= c(0.01,0.01))+
  theme_custom() +
  theme(axis.text.x = element_text(angle=90),
        legend.position = "none")

accmag_post_nadir

```

### 2.1.1 Example passage plot

```{r c_39_plot}

#### Create example passage plot 

C39 <- denver_data %>%
  filter (long_id == "C390327085054")

# # Plot plotly to identify trim times etc
# plot_ly(C39, x = ~time) %>%
#   add_lines(y = ~pres, name = "Pres", line = list(color = "black"), yaxis = "y") %>%
#   add_lines(y = ~accmag, name = "Acc", line = list(color = "red"), yaxis = "y2") %>%
#   layout(
#     xaxis = list(title = "Time (s)"),
#     yaxis = list(title = "Pressure (kPa)", titlefont = list(color = "black")),
#     yaxis2 = list(title = "Acceleration Magnitude (g)", titlefont = list(color = "red"),
#                   overlaying = "y", side = "right")
#   )


# decided to Trim data
# 3s plot
C39 <- C39 %>%
  filter(time >= 66, time <= 69)

# Identify ROI boundaries 
roi_changes <- C39 %>%
  mutate(roi_change = roi != lag(roi, default = first(roi))) %>%
  filter(roi_change) %>%
  pull(time)

#Identify accmag and pressure min max for secondary axis scaling
pres_min <- 90
pres_max <- 120
accmag_min <- min(C39$accmag)  # approximately -5.809
accmag_max <- 230

# Calculate scaling to map accmag range onto pressure range
scale_factor <- (pres_max - pres_min) / (accmag_max - accmag_min)
offset <- pres_min - accmag_min * scale_factor


C39_plot <- ggplot(C39, aes(x = time)) +
  geom_rect(data = C39 %>%
              summarise(xmin = min(time), xmax = max(time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey90", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(data = C39 %>% filter(roi == "Nadir") %>% 
              summarise(xmin = min(time), xmax = max(time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#F4A6A3", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(data = C39 %>% filter(roi == "Post-nadir") %>% 
              summarise(xmin = min(time), xmax = max(time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#A8DADC", alpha = 0.1, inherit.aes = FALSE) +
  geom_line(aes(y = pres), color = "black", size = 0.5) +
  geom_line(aes(y = accmag * scale_factor + offset), color = "red2", size = 0.5) +
  geom_vline(xintercept = roi_changes, color = "blue4", linetype = "dotted") +
  scale_x_continuous(breaks = seq(65, 69, by = 0.25), expand = c(0,0)) +
  scale_y_continuous(
    name = "Pressure (kPa)",
    breaks = seq(90, 120, by = 2),
    limits = c(90, 120),
    sec.axis = sec_axis(
      ~ (. - offset) / scale_factor, 
      name = expression("Acceleration Magnitude (m s" ^-2*")"),
      breaks = seq(0, 230, by = 20)
    )
  ) +
  labs(x = "Time (s)") +
  theme_custom() +
  theme(
    axis.title.y.right = element_text(color = "red2"),
    axis.text.y.right = element_text(color = "red2"),
    axis.text.x = element_text(angle = 90)
  ) +
  coord_cartesian(clip = "off")

C39_plot

# Export svg for editing in Inkscape. Ensure units are cm and size appropriate for page space. No DPI needed (vector)
#ggsave("C39_plot.svg", C39_plot, units = "cm", width = 16, height = 10)

# Export png. Ensure units are cm and size appropriate for page space. Set DPI at 300
#ggsave("C39_plot.png", C39_plot, dpi = 300, units = "cm", width = 16, height = 10)



```

Describe why diagnostics form the plot which lead to further analysis...


# 3. Pressure indicators

Data preparation

```{r pres_prep}

# Filter and shape data accordingly for pressure analysis
#find nadir pressure time and row index 

nadir_tbl <- denver_data %>%
  mutate(row_index = row_number()) %>%          # index in the full table
  filter(roi == "Nadir", !is.na(pres), !is.na(time)) %>%
  group_by(long_id, treatment) %>%
  slice_min(order_by = pres, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    long_id,
    treatment,
    nadir = pres,
    nadir_time = time
  )



#find maximum pressure within 1s before nadir pressure for every sensor per treatment
max_pres_before_nadir <- denver_data %>%
  mutate(row_index = row_number()) %>%
  inner_join(nadir_tbl, by = c("long_id", "treatment")) %>%
  group_by(long_id, treatment, nadir, nadir_time) %>%
  # keep only samples in [nadir_time - 1, nadir_time)
  filter(!is.na(time), !is.na(pres),
         time >= (nadir_time - 1),
         time < nadir_time) %>%
  # take the row with maximum pressure in that window
  slice_max(order_by = pres, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    long_id, treatment,
    nadir, nadir_time, 
    max_p_1s_nadir = pres,
    max_p_1s_nadir_time = time
  )

# calculating rate pressure change (rpc) and log ration pressure change (lrpc)
#setting surface acclimation pressure to 100 kPa (at 0m)
acclim_press <- 100

pressure_tbl <- max_pres_before_nadir %>%
  mutate(
    rate_pc  = max_p_1s_nadir - nadir,
    ratio_pc = acclim_press/ nadir,
    log_rpc  = log(ratio_pc)
  ) %>%
  transmute(long_id, treatment, nadir, rate_pc, ratio_pc, log_rpc)


#filter out extreme outliers
pressure_tbl_2<-pressure_tbl %>%
  filter(!is.na(log_rpc), log_rpc <= 3)%>%
  filter(nadir >= 25)


```

## 3.1 Pressure summary

```{r pres_sum}

# Summarise and tabulate pressure variables across treatments etc

summary_pres_tbl <- pressure_tbl_2 %>%
  group_by(treatment) %>%
  summarise(
    n = n(),
    across(
      c(nadir, rate_pc, ratio_pc, log_rpc),
      list(
        median = ~ median(.x, na.rm = TRUE),
        mean   = ~ mean(.x, na.rm = TRUE),
        Q1     = ~ quantile(.x, 0.25, na.rm = TRUE),
        Q3     = ~ quantile(.x, 0.75, na.rm = TRUE),
        IQR    = ~ IQR(.x, na.rm = TRUE),
        sd     = ~ sd(.x, na.rm = TRUE),
        min    = ~ min(.x, na.rm = TRUE),
        max    = ~ max(.x, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    -treatment,
    names_to = "variable_stat",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = treatment,
    values_from = value
  )

summary_pres_tbl

```

## 3.2 Pressure vis

```{r pres_vis}

# Build core figures here
nadir_plot<-ggplot(pressure_tbl_2, aes(x = treatment, y = nadir)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, #outlier.alpha = 0.8,
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5) +
  geom_hline(yintercept = 50, colour = "red", linetype = "dashed", linewidth = 0.5) +
  annotate("text", x = -Inf, y = Inf, label = "a)", hjust = -0.2, vjust = 1.2, size = 3.5) +
  scale_x_discrete(labels = c(
    "30 RPM Original" = "30",
    "40 RPM Original" = "40",
    "50 RPM Original" = "50",
    "60 RPM Original" = "60",
    "40 RPM Modified" = "40",
    "50 RPM Modified" = "50"
  )) +
  # put group labels roughly under the middle of each block
  annotate("text", x = 2.5, y = -Inf, label = "O", vjust = 2.7, size = 4) +
  annotate("text", x = 5.5, y = -Inf, label = "M", vjust = 2.7, size = 4) +
  theme_minimal() +
  theme_custom() +
   theme(
    #axis.text.x = element_text(angle = 65, hjust = 1),
    axis.text.y.right  = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.title.y.right = element_blank()
  ) +
  labs(x = NULL, y = "Nadir pressure (kPa)") +
  scale_y_continuous(limits =c(50,105),
                     breaks = seq(50, 100, 10)) +
  coord_cartesian( clip = "off")
 
#nadir_plot


rpc_plot <- ggplot(pressure_tbl_2, aes(x = treatment, y = rate_pc)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4,# outlier.alpha = 0.8,
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5) +
  geom_hline(yintercept = 550, colour = "red", linetype = "dashed", linewidth = 0.5) +
  annotate("text", x = -Inf, y = Inf, label = "b)", hjust = -0.2, vjust = 1.2, size = 3.5) +
  scale_x_discrete(labels = c(
    "30 RPM Original" = "30",
    "40 RPM Original" = "40",
    "50 RPM Original" = "50",
    "60 RPM Original" = "60",
    "40 RPM Modified" = "40",
    "50 RPM Modified" = "50"
  )) +
  # put group labels roughly under the middle of each block
  annotate("text", x = 2.5, y = -Inf, label = "O", vjust = 2.7, size = 4) +
  annotate("text", x = 5.5, y = -Inf, label = "M", vjust = 2.7, size = 4) +
    theme_minimal() +
  theme_custom() +
  theme(
    #axis.text.x = element_text(angle = 65, hjust = 1),
    axis.text.y.right  = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.title.y.right = element_blank()
  ) +
  labs(x = NULL, y = expression("Rate pressure change (kPa s"^-1*")")) +
  scale_y_continuous(breaks = seq(0, 560, 50)) +
  #scale_y_break(c(55, 500), scales = 0.6) +
  coord_cartesian(clip = "off")

#rpc_plot


lrpc_plot <- ggplot(pressure_tbl_2, aes(x = treatment, y = log_rpc)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, #outlier.alpha = 0.8,     # "x"
               outlier.size = 0.9 ,outlier.stroke = 0.4, staplewidth = 0.5) +
  geom_hline(yintercept = 0.5, colour = "red", linetype = "dashed",  linewidth = 0.5)+
  #theme_minimal() +
  annotate(
    "text",
    x = -Inf,
    y = Inf,
    label = "c)",
    hjust = -0.2,
    vjust = 1.2,
    size = 3.5
  ) +
  labs(#title = "log_RPC by RPM Scenario",
    x = NULL,
    y = "Log ratio pressure change")+
  scale_x_discrete(labels = c(
     "30 RPM Original" = "30",
    "40 RPM Original" = "40",
    "50 RPM Original" = "50",
    "60 RPM Original" = "60",
    "40 RPM Modified" = "40",
    "50 RPM Modified" = "50"
  )) +
  # put group labels roughly under the middle of each block
  annotate("text", x = 2.5, y = -Inf, label = "O", vjust = 2.7, size = 4) +
  annotate("text", x = 5.5, y = -Inf, label = "M", vjust = 2.7, size = 4) +
  theme_custom() +
 # theme(
    # text = element_text(family = base_family, size = base_size),
    # axis.title = element_text(size = base_size),
    # axis.text  = element_text(size = base_size),
    #axis.text.x = element_text(angle = 65, hjust = 2)
  #)+
  scale_y_continuous(limits = c(0,0.55),
                     breaks = seq(0, 0.55, 0.1)) +
  coord_cartesian( clip = "off")


#lrpc_plot


pres_plots <- plot_grid(nadir_plot, rpc_plot, lrpc_plot,
                        ncol = 3, nrow = 1)

pres_plots

#ggsave(filename = "pres_plots.svg", plot = pres_plots,
#       device = "svg", units = "cm", width = 16, height = 8, dpi = 300)
```

## 3.3 Pressure analysis

```{r pres_analysis}

# Build core analysis here
kruskal.test(nadir ~ treatment,
             data = pressure_tbl_2)

kruskal.test(rate_pc~ treatment,
             data = pressure_tbl_2)

kruskal.test(log_rpc~ treatment,
             data = pressure_tbl_2)
```

# 4. Acceleration indicators

Describe data preparation...

```{r acc_prep}

# Filter and shape data accordingly for acceleration analysis

#filter for rpm 50

max_accmag_data_nadir <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")

plot_data_nadir <- denver_data %>% 
  filter(rpm == 50) %>%
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_nadir <- plot_data_nadir %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


# Get max accmag for each sensor in Post-nadir ROI
max_accmag_data_post <- denver_data %>% 
  filter(rpm == 50) %>%
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")


plot_data_post <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time),
         mean_strain = round(mean_strain)
         ) %>%
  ungroup()

exceed_data_post <- plot_data_post %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


#filter for rpm 40
max_accmag_data_nadir_40 <- denver_data %>%
  filter(rpm == 40) %>% 
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")

plot_data_nadir_40 <- denver_data %>% 
  filter(rpm == 40) %>%
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_nadir_40 <- plot_data_nadir_40 %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


# Get max accmag for each sensor in Post-nadir ROI
max_accmag_data_post_40 <- denver_data %>% 
  filter(rpm == 40) %>%
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")


plot_data_post_40 <- denver_data %>%
  filter(rpm == 40) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_post_40 <- plot_data_post_40 %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()

```

## 4.1 Acceleration summary

Summarise and tabulate acceleration variables across treatments etc

e.g.,central tendencies

```{r acc_sum}

max_accmag_data_post %>%
  group_by(configuration) %>%
  summarise(
    n = n(),
    median = median(max_accmag),
    Q1 = quantile(max_accmag, 0.25),
    Q3 = quantile(max_accmag, 0.75),
    IQR = IQR(max_accmag),
    mean = mean(max_accmag),
    sd = sd(max_accmag),
    .groups = "drop"
  )


```

## 4.2 Acceleration vis

```{r acc_vis}

# Build core figures here
#core figures for 50_O and 50_M

n_labels <- max_accmag_data_nadir %>%
  group_by(configuration) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

max_accmag_boxplot_nadir <- ggplot(max_accmag_data_nadir, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_fill_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")

#max_accmag_boxplot_nadir


exceedance_plot_nadir <- ggplot(exceed_data_nadir, aes(x = exceed_pct, y = accmag, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")

#exceedance_plot_nadir

combined_plot_nadir <-plot_grid(max_accmag_boxplot_nadir, exceedance_plot_nadir,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")
#combined_plot_nadir

max_accmag_boxplot_post <- ggplot(max_accmag_data_post, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "steelblue4",
                                "Modified" = "steelblue2")) +
  scale_fill_manual(values = c("Original" = "steelblue4", 
                               "Modified" = "steelblue2")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
  #coord_cartesian(clip = "off")

#max_accmag_boxplot_post

exceedance_plot_post <- ggplot(exceed_data_post, aes(x = exceed_pct, y = accmag, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
coord_cartesian(clip = "off")

#exceedance_plot_post

combined_plot_post_nadir <-plot_grid(max_accmag_boxplot_post,
                                     exceedance_plot_post,
                          ncol = 2, nrow = 1, rel_widths = c(0.6,0.4),
                          align = "h",axis = "tb")
#combined_plot_post_nadir

#ggsave("combined_plot_post_nadir.svg", combined_plot_post_nadir, units = "cm", width = 16, height = 8)



combined_acc_plot <-plot_grid(combined_plot_nadir, 
                              combined_plot_post_nadir,
                              ncol = 1, nrow = 2,
                              rel_heights =  c(0.465,0.535), align = "h",axis = "tb")
combined_acc_plot


#ggsave("combined_acc_plot.svg",
     #  combined_acc_plot,
      # units = "cm",
      # width = 16,
      # height = 16)


```


```{r acc_vis}
#core figures for 40_O and 40_M

n_labels_40 <- max_accmag_data_nadir_40 %>%
  group_by(configuration) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

max_accmag_boxplot_nadir_40 <- ggplot(max_accmag_data_nadir_40, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_fill_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")

#max_accmag_boxplot_nadir_40


exceedance_plot_nadir_40 <- ggplot(exceed_data_nadir_40, aes(x = exceed_pct, y = accmag, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")



#exceedance_plot_nadir_40

combined_plot_nadir_40 <-plot_grid(max_accmag_boxplot_nadir_40, exceedance_plot_nadir_40,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")
#combined_plot_nadir_40

max_accmag_boxplot_post_40 <- ggplot(max_accmag_data_post_40, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "steelblue4",
                                "Modified" = "steelblue2")) +
  scale_fill_manual(values = c("Original" = "steelblue4", 
                               "Modified" = "steelblue2")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
  #coord_cartesian(clip = "off")

#max_accmag_boxplot_post_40

exceedance_plot_post_40 <- ggplot(exceed_data_post_40, aes(x = exceed_pct, y = accmag, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
coord_cartesian(clip = "off")

#exceedance_plot_post_40

combined_plot_post_nadir_40 <-plot_grid(max_accmag_boxplot_post_40,
                                     exceedance_plot_post_40,
                          ncol = 2, nrow = 1, rel_widths = c(0.6,0.4),
                          align = "h",axis = "tb")

#combined_plot_post_nadir_40

#ggsave("combined_plot_post_nadir.svg", combined_plot_post_nadir, units = "cm", width = 16, height = 8)



combined_acc_plot_40 <-plot_grid(combined_plot_nadir_40, 
                              combined_plot_post_nadir_40,
                              ncol = 1, nrow = 2,
                              rel_heights =  c(0.465,0.535), align = "h",axis = "tb")
combined_acc_plot_40


#ggsave("combined_acc_plot_40.svg",
      # combined_acc_plot_40,
      # units = "cm",
      # width = 16,
      # height = 16)


```

## 4.3 Acceleration analysis

Non-parametric tests for  differences in median acceleration maxima and kernel density .

```{r acc_analysis}

# Core analysis here

#analysis of 50_O vs 50_M
ad.test(
  max_accmag_data_nadir$max_accmag[max_accmag_data_nadir$
                                     configuration == "Original"],
  max_accmag_data_nadir$max_accmag[max_accmag_data_nadir$
                                     configuration == "Modified"]
  )

wilcox.test(max_accmag ~ configuration,
            max_accmag_data_nadir)

ad.test(
  exceed_data_nadir$accmag[exceed_data_nadir$
                             configuration == "Original"],
  exceed_data_nadir$accmag[exceed_data_nadir$
                             configuration == "Modified"]
  )

ad.test(
  max_accmag_data_post$max_accmag[max_accmag_data_post$
                                    configuration == "Original"],
  max_accmag_data_post$max_accmag[max_accmag_data_post$
                                    configuration == "Modified"]
  )

wilcox.test(max_accmag ~ configuration,
            max_accmag_data_post)

ad.test(
  exceed_data_post$accmag[exceed_data_post$
                            configuration == "Original"],
  exceed_data_post$accmag[exceed_data_post$
                            configuration == "Modified"]
  )


```
```{r acc_analysis}


#analysis of 40_O vs 40_M
ad.test(
  max_accmag_data_nadir_40$max_accmag[max_accmag_data_nadir_40$
                                     configuration == "Original"],
  max_accmag_data_nadir_40$max_accmag[max_accmag_data_nadir_40$
                                     configuration == "Modified"]
  )

wilcox.test(max_accmag ~ configuration,
            max_accmag_data_nadir_40)

ad.test(
  exceed_data_nadir_40$accmag[exceed_data_nadir_40$
                             configuration == "Original"],
  exceed_data_nadir_40$accmag[exceed_data_nadir_40$
                             configuration == "Modified"]
  )

ad.test(
  max_accmag_data_post_40$max_accmag[max_accmag_data_post_40$
                                    configuration == "Original"],
  max_accmag_data_post_40$max_accmag[max_accmag_data_post_40$
                                    configuration == "Modified"]
  )

wilcox.test(max_accmag ~ configuration,
            max_accmag_data_post_40)

ad.test(
  exceed_data_post_40$accmag[exceed_data_post_40$
                            configuration == "Original"],
  exceed_data_post_40$accmag[exceed_data_post_40$
                            configuration == "Modified"]
  )



```


## 4.4 Acceleration indicators between 40_M and 50_M
```{r acc_prep}

# Filter and shape data accordingly for acceleration analysis

max_accmag_data_mod_nadir <- denver_data %>%
  filter(rpm %in% c(40,50),
         configuration == "Modified") %>% 
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration, rpm) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")

plot_data_mod_nadir <- denver_data %>% 
  filter(rpm %in% c(40,50),
         configuration == "Modified") %>%
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration, rpm) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_mod_nadir <- plot_data_mod_nadir %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(rpm, configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


# Get max accmag for each sensor in Post-nadir ROI
max_accmag_data_mod_post <- denver_data %>% 
  filter(rpm %in% c(40,50),
         configuration == "Modified") %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration, rpm) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")


plot_data_mod_post <- denver_data %>%
  filter(rpm %in% c(40,50),
         configuration == "Modified") %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration, rpm) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time),
         mean_strain = round(mean_strain)
         ) %>%
  ungroup()

exceed_data_mod_post <- plot_data_mod_post %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(rpm, configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


max_accmag_data_mod_nadir <- max_accmag_data_mod_nadir %>%
  dplyr::mutate(rpm = factor(rpm, levels = c(40, 50), labels = c("40", "50")))

exceed_data_mod_nadir <- exceed_data_mod_nadir %>%
  dplyr::mutate(rpm = factor(rpm, levels = c(40, 50), labels = c("40", "50")))

max_accmag_data_mod_post <- max_accmag_data_mod_post %>%
  dplyr::mutate(rpm = factor(rpm, levels = c(40, 50), labels = c("40", "50")))

exceed_data_mod_post <- exceed_data_mod_post %>%
  dplyr::mutate(rpm = factor(rpm, levels = c(40, 50), labels = c("40", "50")))



```

## 4.5 Acceleration 40_M vs 50_M vis

```{r acc_40_50_mod_vis}

# Build core figures here

n_mod_labels <- max_accmag_data_mod_nadir %>%
  group_by(rpm) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

max_accmag_boxplot_mod_nadir <- ggplot(max_accmag_data_mod_nadir, aes(y = max_accmag))+
  stat_halfeye(
    aes(fill = rpm),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = rpm),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = rpm,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("40" = "red1", "50" = "pink4")) +
  scale_fill_manual(values = c("40" = "red1", "50" = "pink4")) +
  geom_text(
    data = n_mod_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~rpm,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(rpm = c(
      "40" = "40 RPM Modified Configuration",
      "50" = "50 RPM Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")


#max_accmag_boxplot_mod_nadir


exceedance_plot_mod_nadir <- ggplot(exceed_data_mod_nadir, aes(x = exceed_pct, y = accmag, group = rpm)) +
  geom_line(aes(color = rpm, linetype = rpm), linewidth = 1) +
  scale_color_manual(values = c("40" = "red1", "50" = "pink4")) +
  scale_linetype_manual(values = c("40" = "solid", "50" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")

#exceedance_plot_nadir

combined_plot_mod_nadir <-plot_grid(max_accmag_boxplot_mod_nadir, exceedance_plot_mod_nadir,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")
#combined_plot_mod_nadir

max_accmag_boxplot_mod_post <- ggplot(max_accmag_data_mod_post, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = rpm),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = rpm),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = rpm,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("40" = "steelblue4",
                                "50" = "steelblue2")) +
  scale_fill_manual(values = c("40" = "steelblue4", 
                               "50" = "steelblue2")) +
  geom_text(
    data = n_mod_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~rpm,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(rpm = c(
      "40" = "40 RPM Modified Configuration",
      "50" = "50 RPM Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
  #coord_cartesian(clip = "off")

#max_accmag_boxplot_mod_post

exceedance_plot_mod_post <- ggplot(exceed_data_mod_post, aes(x = exceed_pct, y = accmag, group = rpm)) +
  geom_line(aes(color = rpm, linetype = rpm), linewidth = 1) +
  scale_color_manual(values = c("40" = "steelblue4", "50" = "steelblue2")) +
  scale_linetype_manual(values = c("40" = "solid", "50" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
coord_cartesian(clip = "off")

#exceedance_plot_mod_post

combined_plot_mod_post_nadir <-plot_grid(max_accmag_boxplot_mod_post,
                                     exceedance_plot_mod_post,
                          ncol = 2, nrow = 1, rel_widths = c(0.6,0.4),
                          align = "h",axis = "tb")
#combined_plot_post_nadir

#ggsave("combined_plot_post_nadir.svg", combined_plot_post_nadir, units = "cm", width = 16, height = 8)



combined_acc_plot_mod <-plot_grid(combined_plot_mod_nadir, 
                              combined_plot_mod_post_nadir,
                              ncol = 1, nrow = 2,
                              rel_heights =  c(0.465,0.535), align = "h",axis = "tb")
combined_acc_plot_mod


#ggsave("combined_acc_plot_mod.svg",
     #  combined_acc_plot_mod,
     # units = "cm",
     #  width = 16,
      # height = 16)

```


## 4.6 Acceleration analysis of 40_M vs 50_M

Non-parametric tests for  differences in median acceleration maxima and kernel density .

```{r acc_analysis}

# Core analysis here

ad.test(
  max_accmag_data_mod_nadir$max_accmag[max_accmag_data_mod_nadir$
                                     rpm == "40"],
  max_accmag_data_mod_nadir$max_accmag[max_accmag_data_mod_nadir$
                                     rpm == "50"]
  )

wilcox.test(max_accmag ~ rpm,
            max_accmag_data_mod_nadir)

ad.test(
  exceed_data_mod_nadir$accmag[exceed_data_mod_nadir$
                             rpm == "40"],
  exceed_data_mod_nadir$accmag[exceed_data_mod_nadir$
                             rpm == "50"]
  )

ad.test(
  max_accmag_data_mod_post$max_accmag[max_accmag_data_mod_post$
                                    rpm == "40"],
  max_accmag_data_mod_post$max_accmag[max_accmag_data_mod_post$
                                    rpm == "50"]
  )

wilcox.test(max_accmag ~ rpm,
            max_accmag_data_mod_post)

ad.test(
  exceed_data_mod_post$accmag[exceed_data_mod_post$
                            rpm == "40"],
  exceed_data_mod_post$accmag[exceed_data_mod_post$
                            rpm == "50"]
  )


```


# 5. Strain indicators

Data preparation

```{r strain_prep}

# Filter and shape data accordingly for Strain analysis
#filter for nadir and post-nadir region for general boxplots of all treatment groups

strain_data_nadir <- denver_data %>%
  #filter(rpm == 50) %>% 
  filter(roi == "Nadir") %>%
   group_by(long_id, treatment, configuration) %>%
  summarise(mean_mean_strain = mean(mean_strain), .groups = "drop")

strain_data_post_nadir <- denver_data %>%
  #filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
   group_by(long_id, treatment, configuration) %>%
  summarise(mean_mean_strain = mean(mean_strain), .groups = "drop")



#comparison of 50_M vs 50_O data preparation mean strain

strain_data_nadir_50 <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(mean_mean_strain = mean(mean_strain), .groups = "drop")

plot_data_nadir <- denver_data %>% 
  filter(rpm == 50) %>%
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_nadir_strain_50 <- plot_data_nadir %>%
  #filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(mean_strain, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(mean_strain),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


strain_data_post_nadir_50 <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(mean_mean_strain = mean(mean_strain), .groups = "drop")



exceed_data_post_nadir_strain_50 <- plot_data_post %>%
  #filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(mean_strain, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(mean_strain),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()



#comparison of 50_M vs 50_O data preparation max strain

max_strain_data_nadir_50 <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_mean_strain = max(mean_strain), .groups = "drop")

plot_data_nadir <- denver_data %>% 
  filter(rpm == 50) %>%
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_nadir_strain_50 <- plot_data_nadir %>%
  #filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(mean_strain, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(mean_strain),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


max_strain_data_post_nadir_50 <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_mean_strain = max(mean_strain), .groups = "drop")



exceed_data_post_nadir_strain_50 <- plot_data_post %>%
  #filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(mean_strain, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(mean_strain),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()
```



## 5.1 Strain summary

```{r strain_sum}

# Summarise and tabulate Strain variables across treatments etc


strain_data_nadir_50 %>%
  group_by(configuration) %>%
  summarise(
    n = n(),
    median = median(mean_mean_strain),
    Q1 = quantile(mean_mean_strain, 0.25),
    Q3 = quantile(mean_mean_strain, 0.75),
    IQR = IQR(mean_mean_strain),
    mean = mean(mean_mean_strain),
    sd = sd(mean_mean_strain),
    min= min(mean_mean_strain),
    max = max(mean_mean_strain),
    .groups = "drop"
  )



```

## 5.2 Strain vis

```{r strain_vis}

# Build core figures here

#all treatments boxplot nadir region
mean_strain_nadir_plot<-ggplot(strain_data_nadir, aes(x = treatment, y = mean_mean_strain, color = configuration)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, #outlier.alpha = 0.8,
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5) +
  #geom_hline(yintercept = 500, colour = "red", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = -Inf, y = Inf, label = "a)", hjust = -0.2, vjust = 1.2, size = 3.5) #+
  scale_x_discrete(labels = c(
    "30 RPM Original" = "30",
    "40 RPM Original" = "40",
    "50 RPM Original" = "50",
    "60 RPM Original" = "60",
    "40 RPM Modified" = "40",
    "50 RPM Modified" = "50"
  )) +
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  #scale_color_manual(values = c(Original = "red1", Modified = "pink4"))+
  # put group labels roughly under the middle of each block
  annotate("text", x = 2.5, y = -Inf, label = "O", vjust = 2.7, size = 4) +
  annotate("text", x = 5.5, y = -Inf, label = "M", vjust = 2.7, size = 4) +
  theme_minimal() +
  theme_custom() +
   theme(
    #axis.text.x = element_text(angle = 65, hjust = 1),
    axis.text.y.right  = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.title.y.right = element_blank()
  ) +
  labs(x = NULL, y = "Mean strain rate (l s-1)") +
  scale_y_continuous(limits =c(0,5000),
                     breaks = seq(0, 5000, 500)) +
  coord_cartesian( clip = "off")

#mean_strain_nadir_plot


#all treatments boxplots post-nadir region
mean_strain_post_nadir_plot<-ggplot(strain_data_post_nadir, aes(x = treatment, y = mean_mean_strain, color = configuration)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, #outlier.alpha = 0.8,
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5) +
  #geom_hline(yintercept = 500, colour = "red", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = -Inf, y = Inf, label = "a)", hjust = -0.2, vjust = 1.2, size = 3.5) #+
  scale_x_discrete(labels = c(
    "30 RPM Original" = "30",
    "40 RPM Original" = "40",
    "50 RPM Original" = "50",
    "60 RPM Original" = "60",
    "40 RPM Modified" = "40",
    "50 RPM Modified" = "50"
  )) +
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  # put group labels roughly under the middle of each block
  annotate("text", x = 2.5, y = -Inf, label = "O", vjust = 2.7, size = 4) +
  annotate("text", x = 5.5, y = -Inf, label = "M", vjust = 2.7, size = 4) +
  theme_minimal() +
  theme_custom() +
   theme(
    #axis.text.x = element_text(angle = 65, hjust = 1),
    axis.text.y.right  = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.title.y.right = element_blank()
  ) +
  labs(x = NULL, y = "Mean strain rate (l s-1)") +
  scale_y_continuous(limits =c(0,5000),
                     breaks = seq(0, 5000, 500)) +
  coord_cartesian( clip = "off")

#mean_strain_post_nadir_plot




combined_boxplot_strain<-plot_grid(mean_strain_nadir_plot, mean_strain_post_nadir_plot,
                                ncol = 2, nrow = 1, rel_widths = c(0.5,0.5), align = "h",axis = "tb")


combined_boxplot_strain



#ggsave("combined_boxplot_strain.svg",
     #  combined_boxplot_strain,
     # units = "cm",
     #  width = 16,
      # height = 16)


```


```{r strain_vis}
# figures for comparison of mean strain 50_M vs 50_O

n_labels <- strain_data_nadir_50 %>%
  group_by(configuration) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

strain_boxplot_nadir <- ggplot(strain_data_nadir_50, aes(y = mean_mean_strain)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_fill_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 4000)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Mean strain rate (l s"^-1*")")
  ) +
  scale_y_continuous(breaks = seq(0, 5000, 500), limits = c(0,5000)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")

#strain_boxplot_nadir

exceedance_plot_nadir_strain <- ggplot(exceed_data_nadir_strain_50, aes(x = exceed_pct, y = mean_strain, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 5000, 500), limits = c(0, 5000)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Mean strain rate (l s"^-1*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")

#exceedance_plot_nadir_strain

combined_plot_nadir_strain <-plot_grid(strain_boxplot_nadir, exceedance_plot_nadir_strain,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")


strain_boxplot_post_nadir <- ggplot(strain_data_post_nadir_50, aes(y = mean_mean_strain)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_fill_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 4000)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Mean strain rate (l s"^-1*")")
  ) +
  scale_y_continuous(breaks = seq(0, 5000, 500), limits = c(0,5000)) +
  theme_custom() +
   theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
#coord_cartesian(clip = "off")


exceedance_plot_post_nadir_strain <- ggplot(exceed_data_post_nadir_strain_50, aes(x = exceed_pct, y = mean_strain, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 5000, 500), limits = c(0, 5000)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Mean strain rate (l s"^-1*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
  coord_cartesian(clip = "off")



combined_plot_post_nadir_strain <-plot_grid(strain_boxplot_post_nadir, exceedance_plot_post_nadir_strain,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")



combined_strain_plot_50 <-plot_grid(combined_plot_nadir_strain , 
                              combined_plot_post_nadir_strain ,
                              ncol = 1, nrow = 2,
                              rel_heights =  c(0.465,0.535), align = "h",axis = "tb")

combined_strain_plot_50


#ggsave("combined_strain_plot_50.svg",
     #  combined_strain_plot_50,
     # units = "cm",
     #  width = 16,
      # height = 16)

```
```{r strain_vis}
# figures for comparison of max strain 50_M vs 50_O

n_labels <- strain_data_nadir_50 %>%
  group_by(configuration) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

max_strain_boxplot_nadir <- ggplot(max_strain_data_nadir_50, aes(y = max_mean_strain)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_fill_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 25000)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Mean strain rate (l s"^-1*")")
  ) +
  scale_y_continuous(breaks = seq(0, 30000, 5000), limits = c(0,30000)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")

#max_strain_boxplot_nadir

exceedance_plot_nadir_max_strain <- ggplot(exceed_data_nadir_strain_50, aes(x = exceed_pct, y = mean_strain, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 30000, 5000), limits = c(0, 30000)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Mean strain rate (l s"^-1*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")

#exceedance_plot_nadir_strain

combined_plot_nadir_max_strain <-plot_grid(max_strain_boxplot_nadir, exceedance_plot_nadir_max_strain,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")


max_strain_boxplot_post_nadir <- ggplot(max_strain_data_post_nadir_50, aes(y = max_mean_strain)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_fill_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 25000)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Mean strain rate (l s"^-1*")")
  ) +
  scale_y_continuous(breaks = seq(0, 30000, 5000), limits = c(0,30000)) +
  theme_custom() +
   theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
#coord_cartesian(clip = "off")


exceedance_plot_post_nadir_max_strain <- ggplot(exceed_data_post_nadir_strain_50, aes(x = exceed_pct, y = mean_strain, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 30000, 5000), limits = c(0, 30000)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Mean strain rate (l s"^-1*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
  coord_cartesian(clip = "off")



combined_plot_post_nadir_max_strain <-plot_grid(max_strain_boxplot_post_nadir, exceedance_plot_post_nadir_max_strain,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")



combined_max_strain_plot_50 <-plot_grid(combined_plot_nadir_max_strain , 
                              combined_plot_post_nadir_max_strain ,
                              ncol = 1, nrow = 2,
                              rel_heights =  c(0.465,0.535), align = "h",axis = "tb")

combined_max_strain_plot_50
```



## 5.3 Strain analysis

```{r strain_analysis}

# testing differences between mean_strain in nadir region between different operational scenarios
kruskal.test(mean_mean_strain ~ treatment,
             data = strain_data_nadir)


# testing differences between mean_strain in post-nadir region between different operational scenarios
kruskal.test(mean_mean_strain ~ treatment,
             data = strain_data_post_nadir)


#testing differences in nadir region between mean strain in 50_M and 50_O
ad.test(
  strain_data_nadir_50$mean_mean_strain[strain_data_nadir_50$
                                     configuration == "Original"],
 strain_data_nadir_50$mean_mean_strain[strain_data_nadir_50$
                                     configuration== "Modified"]
  )

wilcox.test(mean_mean_strain ~ configuration,
            strain_data_nadir_50)



#testing differences in post-nadir region between mean strain in 50_M and 50_O
ad.test(
  strain_data_post_nadir_50$mean_mean_strain[strain_data_post_nadir_50$
                                     configuration == "Original"],
 strain_data_post_nadir_50$mean_mean_strain[strain_data_post_nadir_50$
                                     configuration== "Modified"]
  )

wilcox.test(mean_mean_strain ~ configuration,
           strain_data_post_nadir_50)

```

# 6. Fish survival estimates

Calculate 95% 'Wilson binomial confidence intervals and build visualization.

```{r fish_survival}

survival_data <- tibble(
  species = c(
    "Roach (Rutilus rutilus)",
    "European eel (Anguilla anguilla) - adult",
    "European eel (Anguilla anguilla) - juvenile",
    "Bream (Abramis brama)",
    "Brown trout (Salmo trutta)"
  ),
  total_recap = c(239, 46, 219, 208, 211),
  total_surv  = c(237, 46, 219, 207, 211),
  group       = NA,
  survival    = NA,
  lower_ci    = NA,
  upper_ci    = NA
)

survival_ci <- survival_data %>%
  mutate(
    survival = total_surv / total_recap,
    # Calculate Wilson CIs
    ci_result = map2(total_surv, total_recap, 
                     ~binom.confint(.x, .y, methods = "wilson"))
  ) %>%
  unnest(ci_result) %>%
  select(species, group, total_recap, total_surv, 
         survival = mean, lower_ci = lower, upper_ci = upper)

survival_ci %>% 
  select(-group)

```

```{r fish_survival_plot}

impact_data <- survival_ci %>%
  mutate(species = factor(species, levels = rev(unique(species))))

survival_plot <- ggplot(impact_data, aes(x = survival * 100, y = species)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, 
           fill = "#ffcccc", alpha = 0.3) +
  annotate("rect", xmin = 96, xmax = 100, ymin = -Inf, ymax = Inf, 
           fill = "#ccffcc", alpha = 0.5) +
  geom_vline(xintercept = c(96, 98, 100), linetype = "dashed", 
             color = "grey60", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = lower_ci * 100, xmax = upper_ci * 100), 
                 height = 0.2, linewidth = 0.5) +
  geom_point(size = 2, shape = 16) +
  geom_text(aes(label = paste0("italic(n) == ", total_recap)), 
            x = 90.5, hjust = 0, size = 3.5, parse = TRUE) +
  annotate("text", x = 90, y = Inf, 
           label = "CI = 95% (Wilson Binomial)\nMargin = 2%",
           hjust = 0, vjust = 1, size = 3.5, lineheight = 0.9) +
  scale_x_continuous(limits = c(90, 100), 
                     breaks = seq(90, 100, 1)) +
  labs(x = "Survival Probability (%)",
       y = NULL) +
  theme_custom() +
  coord_cartesian(clip = "off")

survival_plot

# ggsave("survival_plot.svg", 
#        plot = survival_plot, 
#        width = 18, 
#        height = 10, 
#        units = "cm")

```
# 7. Import backpack data

```{r backpack_data_import}

backpack_files <- list.files(
  "./input",
  pattern = "^backpack_.*\\.csv$",
  full.names = TRUE
)

backpack_data <- backpack_files %>%
  map_dfr(read_csv)

#import extra 50_M BDS data with calculated zero-crossings
BDS_50rpm_modified <- list.files(
  "./input",
  pattern = "^BDS_.*\\.csv$",
  full.names = TRUE
)

BDS_50rpm_data <- BDS_50rpm_modified %>%
  map_dfr(read_csv)


```

##7.1 Visualize Example passage plot of backpack data

```{r EEL04_plot}

#### Create example passage plot 

EEL04 <- backpack_data %>%
  filter (long_id == "EEL040129130045")

# # Plot plotly to identify trim times etc
# plot_ly(EEL04, x = ~Time) %>%
#   add_lines(y = ~pres, name = "Pres", line = list(color = "black"), yaxis = "y") %>%
#   add_lines(y = ~accmag, name = "Acc", line = list(color = "red"), yaxis = "y2") %>%
#   layout(
#     xaxis = list(title = "Time (s)"),
#     yaxis = list(title = "Pressure (kPa)", titlefont = list(color = "black")),
#     yaxis2 = list(title = "Acceleration Magnitude (g)", titlefont = list(color = "red"),
#                   overlaying = "y", side = "right")
#   )


# decided to Trim data
# 3s plot
EEL04 <-EEL04 %>%
  filter(Time >= 15.561, Time <=18.471)

# Identify ROI boundaries 
roi_changes_eel <- EEL04 %>%
  mutate(roi_change = roi != lag(roi, default = first(roi))) %>%
  filter(roi_change) %>%
  pull(Time)

#Identify accmag and pressure min max for secondary axis scaling
pres_min_eel <- 90.644
pres_max_eel <- 109.085
accmag_min_eel <- min(EEL04$accmag)  
accmag_max_eel <- 67.81396

# Calculate scaling to map accmag range onto pressure range
scale_factor_eel <- (pres_max_eel - pres_min_eel) / (accmag_max_eel - accmag_min_eel)
offset_eel <- pres_min_eel - accmag_min_eel * scale_factor_eel


EEL04_plot <- ggplot(EEL04, aes(x = Time)) +
  geom_rect(data = EEL04 %>%
              summarise(xmin = min(Time), xmax = max(Time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey90", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(data = EEL04 %>% filter(roi == "Nadir") %>% 
              summarise(xmin = min(Time), xmax = max(Time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#F4A6A3", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(data = EEL04 %>% filter(roi == "Post-Nadir") %>% 
              summarise(xmin = min(Time), xmax = max(Time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#A8DADC", alpha = 0.1, inherit.aes = FALSE) +
  geom_line(aes(y = pres), color = "black", size = 0.5) +
  geom_line(aes(y = accmag * scale_factor_eel + offset_eel), color = "red2", size = 0.5) +
  geom_vline(xintercept = roi_changes_eel, color = "blue4", linetype = "dotted") +
  scale_x_continuous(breaks = seq(15, 19, by = 0.25), expand = c(0,0)) +
  scale_y_continuous(
    name = "Pressure (kPa)",
    breaks = seq(90, 110, by = 2),
    limits = c(90, 110),
    sec.axis = sec_axis(
      ~ (. - offset) / scale_factor, 
      name = expression("Acceleration Magnitude (m s" ^-2*")"),
      breaks = seq(0, 110, by = 10)
    )
  ) +
  labs(x = "Time (s)") +
  theme_custom() +
  theme(
    axis.title.y.right = element_text(color = "red2"),
    axis.text.y.right = element_text(color = "red2"),
    axis.text.x = element_text(angle = 90)
  ) +
  coord_cartesian(clip = "off")

EEL04_plot

```

##7.2 backpack pressure data preperation

```{r backpack_pres_prep}



#find nadir pressure time and row 

nadir_tbl_eel <- backpack_data %>%
  mutate(row_index = row_number()) %>%          # index in the full table
  filter(roi == "Nadir", !is.na(pres), !is.na(Time)) %>%
  group_by(long_id, treatment) %>%
  slice_min(order_by = pres, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    long_id,
    treatment,
    nadir = pres,
    nadir_time = Time
  )



#find maximum pressure within 1s before nadir pressure for every sensor per treatment
max_pres_before_nadir_eel <- backpack_data %>%
  mutate(row_index = row_number()) %>%
  inner_join(nadir_tbl_eel, by = c("long_id", "treatment")) %>%
  group_by(long_id, treatment, nadir, nadir_time) %>%
  # keep only samples in [nadir_time - 1, nadir_time)
  filter(!is.na(Time), !is.na(pres),
         Time >= (nadir_time - 1),
         Time < nadir_time) %>%
  # take the row with maximum pressure in that window
  slice_max(order_by = pres, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    long_id, treatment,
    nadir, nadir_time, 
    max_p_1s_nadir = pres,
    max_p_1s_nadir_time = Time
  )

# calculating rate pressure change (rpc) and log ration pressure change (lrpc)
#setting surface acclimation pressure to 100 kPa (at 0m)
acclim_press <- 100

pressure_tbl_eel <- max_pres_before_nadir_eel %>%
  mutate(
    rate_pc  = max_p_1s_nadir - nadir,
    ratio_pc = acclim_press/ nadir,
    log_rpc  = log(ratio_pc)
  ) %>%
  transmute(long_id, treatment, nadir, rate_pc, ratio_pc, log_rpc)



```


##7.3 backpack pressure data summary

```{r backpack_pres_sum}

# Summarise and tabulate pressure variables across treatments etc

summary_pres_tbl_eel <- pressure_tbl_eel %>%
  group_by(treatment) %>%
  summarise(
    n = n(),
    across(
      c(nadir, rate_pc, ratio_pc, log_rpc),
      list(
        median = ~ median(.x, na.rm = TRUE),
        mean   = ~ mean(.x, na.rm = TRUE),
        Q1     = ~ quantile(.x, 0.25, na.rm = TRUE),
        Q3     = ~ quantile(.x, 0.75, na.rm = TRUE),
        IQR    = ~ IQR(.x, na.rm = TRUE),
        sd     = ~ sd(.x, na.rm = TRUE),
        min    = ~ min(.x, na.rm = TRUE),
        max    = ~ max(.x, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    -treatment,
    names_to = "variable_stat",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = treatment,
    values_from = value
  )

summary_pres_tbl_eel

```


##7.4 backpack pressure vis
```{r bakpack_pres_vis}

# Build core figures here
nadir_plot_eel<-ggplot(pressure_tbl_eel, aes(x = treatment, y = nadir)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, #outlier.alpha = 0.8,
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5) +
  geom_hline(yintercept = 50, colour = "red", linetype = "dashed", linewidth = 0.5) +
  annotate("text", x = -Inf, y = Inf, label = "a)", hjust = -0.2, vjust = 1.2, size = 3.5) +
  theme_minimal() +
  theme_custom() +
   theme(
    #axis.text.x = element_text(angle = 65, hjust = 1),
    axis.text.y.right  = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.title.y.right = element_blank()
  ) +
  labs(x = NULL, y = "Nadir pressure (kPa)") +
  scale_y_continuous(limits =c(50,105),
                     breaks = seq(50, 100, 10)) +
  coord_cartesian( clip = "off")
 
#nadir_plot_eel


rpc_plot_eel <- ggplot(pressure_tbl_eel, aes(x = treatment, y = rate_pc)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4,# outlier.alpha = 0.8,
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5) +
  geom_hline(yintercept = 550, colour = "red", linetype = "dashed", linewidth = 0.5) +
  annotate("text", x = -Inf, y = Inf, label = "b)", hjust = -0.2, vjust = 1.2, size = 3.5) +
    theme_minimal() +
  theme_custom() +
  theme(
    #axis.text.x = element_text(angle = 65, hjust = 1),
    axis.text.y.right  = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.title.y.right = element_blank()
  ) +
  labs(x = NULL, y = expression("Rate pressure change (kPa s"^-1*")")) +
  scale_y_continuous(breaks = seq(0, 560, 50)) +
  #scale_y_break(c(55, 500), scales = 0.6) +
  coord_cartesian(clip = "off")

#rpc_plot_eel


lrpc_plot_eel <- ggplot(pressure_tbl_eel, aes(x = treatment, y = log_rpc)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, #outlier.alpha = 0.8,     # "x"
               outlier.size = 0.9 ,outlier.stroke = 0.4, staplewidth = 0.5) +
  geom_hline(yintercept = 0.5, colour = "red", linetype = "dashed",  linewidth = 0.5)+
  #theme_minimal() +
  annotate(
    "text",
    x = -Inf,
    y = Inf,
    label = "c)",
    hjust = -0.2,
    vjust = 1.2,
    size = 3.5
  ) +
  labs(#title = "log_RPC by RPM Scenario",
    x = NULL,
    y = "Log ratio pressure change")+
  theme_custom() +
  scale_y_continuous(limits = c(0,0.55),
                     breaks = seq(0, 0.55, 0.1)) +
  coord_cartesian( clip = "off")


#lrpc_plot_eel


pres_plots_eel <- plot_grid(nadir_plot_eel, rpc_plot_eel, lrpc_plot_eel,
                        ncol = 3, nrow = 1)

pres_plots_eel

#ggsave(filename = "pres_plots_eel.svg", plot = pres_plots_eel,
#       device = "svg", units = "cm", width = 16, height = 8, dpi = 300)
```


##7.6 passage duration in region 1 

```{r backpack_pres_duration}

backpack_roi1 <- backpack_data %>%
  group_by(long_id) %>%
  mutate(
    first_nadir_time = min(Time[roi == "Nadir"], na.rm = TRUE)
  ) %>%
  filter(
    roi == "unlabelled",
    Time < first_nadir_time
  ) %>%
  ungroup()


BDS_50rpm_roi1 <- BDS_50rpm_data %>%
  group_by(long_id) %>%
  mutate(
    first_nadir_time = min(time[roi == "nadir"], na.rm = TRUE)
  ) %>%
  filter(
    roi == "unlabeled",
    time < first_nadir_time
  ) %>%
  ungroup()


passage_duration_r1 <- backpack_roi1 %>%
  group_by(long_id) %>%
  summarise(
    start_time = min(Time, na.rm = TRUE),
    end_time   = max(Time, na.rm = TRUE),
    duration_r1 = end_time - start_time,
    sensor= "backpack",
    .groups = "drop"
  )


passage_duration_BDS_r1 <- BDS_50rpm_roi1 %>%
  group_by(long_id) %>%
  summarise(
    start_time = min(time, na.rm = TRUE),
    end_time   = max(time, na.rm = TRUE),
    duration_r1 = end_time - start_time,
    sensor= "BDS",
    .groups = "drop"
  )

passage_duration <- bind_rows(passage_duration_r1, passage_duration_BDS_r1)

zero_crossings_roi1 <- backpack_roi1 %>%
  group_by(long_id) %>%
  summarise(
    zero_crossings_X = sum(zeroX == 1, na.rm = TRUE),
    sensor= "backpack",
    .groups = "drop"
  )


zero_crossings_BDS_roi1 <- BDS_50rpm_roi1 %>%
  group_by(long_id) %>%
  summarise(
    zero_crossings_X = sum(zeroX == 1, na.rm = TRUE),
    sensor= "BDS",
    .groups = "drop"
  )


zero_crossings <- bind_rows(zero_crossings_BDS_roi1, zero_crossings_roi1)

ODBA_roi1 <- backpack_roi1 %>%
   group_by(long_id) %>%
  summarise(mean_ODBA = mean(ODBA, na.rm = TRUE),
            sensor= "backpack", .groups = "drop")


ODBA_BDS_roi1 <- BDS_50rpm_roi1 %>%
   group_by(long_id) %>%
  summarise(mean_ODBA = mean(ODBA, na.rm = TRUE),
            sensor = "BDS", .groups = "drop")


ODBA <- bind_rows(ODBA_BDS_roi1, ODBA_roi1)


backpack_data <- backpack_data %>%
  mutate(sensor = "backpack")%>%
  rename(time =Time)

BDS_50rpm_data <- BDS_50rpm_data%>%
  mutate(sensor = "BDS")

sensor_data <- bind_rows(backpack_data, BDS_50rpm_data)

```

##7.7 backpack summary

```{r backpack_sum}

passage_duration_r1 %>%
    summarise(
    n = n(),
    median = median(duration_r1),
    Q1 = quantile(duration_r1, 0.25),
    Q3 = quantile(duration_r1, 0.75),
    IQR = IQR(duration_r1),
    mean = mean(duration_r1),
    sd = sd(duration_r1),
    .groups = "drop"
  )

zero_crossings_roi1 %>%
    summarise(
    n = n(),
    median = median(zero_crossings_X),
    Q1 = quantile(zero_crossings_X, 0.25),
    Q3 = quantile(zero_crossings_X, 0.75),
    IQR = IQR(zero_crossings_X),
    mean = mean(zero_crossings_X),
    sd = sd(zero_crossings_X),
    .groups = "drop"
  )


ODBA_roi1 %>%
    summarise(
    n = n(),
    median = median(mean_ODBA),
    Q1 = quantile(mean_ODBA, 0.25),
    Q3 = quantile(mean_ODBA, 0.75),
    IQR = IQR(mean_ODBA),
    mean = mean(mean_ODBA),
    sd = sd(mean_ODBA),
    .groups = "drop"
  )

```
##7.8 backpack vis
```{r backpack_vis_duraion}

duration_boxplot <- ggplot(passage_duration_r1,
       aes(x = "", y = duration_r1)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, 
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5, color ="black") +
   labs(x = NULL,
       y = "Passage duration in Region 1 (s)") +
  theme_custom()+
  theme(legend.position = "none")

#duration_boxplot


zero_crossings_boxplot <- ggplot(zero_crossings_roi1,
       aes(x = "", y = zero_crossings_X)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, 
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5, color ="black") +
  labs(x = NULL,
       y = "Number of zero-crossings (rotation) in Region 1") +
  theme_custom()+
  theme(legend.position = "none")

#zero_crossings_boxplot


mean_ODBA_boxplot <- ggplot(ODBA_roi1,
       aes(x = "", y = mean_ODBA)) +
   stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, 
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5, color ="black") +
  labs(x = NULL,
       y = expression("mean ODBA (m s"^-2*") in Region 1")) +
  theme_custom()+
  theme(legend.position = "none")

#mean_ODBA_boxplot


backpack_vis <- plot_grid(duration_boxplot, zero_crossings_boxplot, mean_ODBA_boxplot,
                        ncol = 3, nrow = 1)

backpack_vis



#ggsave("backpack_vis.svg",
     #  backpack_vis,
     # units = "cm",
     #  width = 16,
      # height = 16)
```


##7.9 backpack vs BDS vis ODBA, passage duration, zero-crossings in region 1

```{r backpack_comparison_vis}

duration_boxplot_2 <- ggplot(passage_duration,
       aes(x = sensor, y = duration_r1, fill = sensor)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, 
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5, color ="black") +
   scale_fill_manual(values = c(
    "BDS" = "grey70",
    "backpack" = "white"
  )) +
  labs(x = NULL,
       y = "Passage duration in Region 1 (s)") +
  theme_custom()+
  theme(legend.position = "none")

#duration_boxplot_2


zero_crossings_boxplot_2 <- ggplot(zero_crossings,
       aes(x = sensor, y = zero_crossings_X, fill = sensor)) +
  stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, 
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5, color ="black") +
   scale_fill_manual(values = c(
    "BDS" = "grey70",
    "backpack" = "white"
  )) +
  labs(x = NULL,
       y = "Number of zero-crossings (rotation) in Region 1") +
  theme_custom()+
  theme(legend.position = "none")

#zero_crossings_boxplot_2


mean_ODBA_boxplot_2 <- ggplot(ODBA_roi1,
       aes(x = sensor, y = mean_ODBA)) +
   stat_boxplot(geom = "errorbar", width = 0.3, position = position_dodge(0.9)) +
  geom_boxplot(outlier.shape = 4, 
               outlier.size = 0.9, outlier.stroke = 0.4, staplewidth = 0.5, color ="black") +
  labs(x = NULL,
       y = expression("mean ODBA (m s"^-2*") in Region 1")) +
  theme_custom()+
  theme(legend.position = "none")

#mean_ODBA_boxplot_2


backpack_vis_2 <- plot_grid(duration_boxplot_2, zero_crossings_boxplot_2, 
                        ncol = 2, nrow = 1)

backpack_vis_2



#ggsave("backpack_vis_2.svg",
     #  backpack_vis_2,
     # units = "cm",
     #  width = 16,
      # height = 16)
```
##7.10 backpack comparison BDS data analysis

```{r backpack_vs_BDs_analysis}
#consider the difference in sample size
#backpack n =14
#BDS n=115

wilcox.test(duration_r1 ~ sensor,
            passage_duration)
            
wilcox.test(zero_crossings_X ~ sensor,
            zero_crossings)
```

##7.11 backpack acceleration vis preparation
```{r backpack_vs_BDs_acc_prep}

# Filter and shape data accordingly for acceleration analysis

#filter for nadir

max_accmag_bp_nadir <- sensor_data %>%
  filter(roi %in% c("nadir", "Nadir"))%>%
  group_by(long_id, sensor) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")

plot_bp_nadir <- sensor_data %>% 
  filter(roi %in% c("nadir", "Nadir"))%>%
  group_by(long_id, sensor) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_bp_nadir <- plot_bp_nadir %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(sensor) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


# Get max accmag for each sensor in Post-nadir ROI
max_accmag_bp_post <- sensor_data %>% 
   filter(roi %in% c("1s_post", "Post-Nadir"))%>%
  group_by(long_id, sensor) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")


plot_bp_post <- sensor_data %>%
   filter(roi %in% c("1s_post", "Post-Nadir"))%>%
  group_by(long_id, sensor) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_bp_post <- plot_bp_post %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(sensor) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


```

##7.12 backpack acceleration vis
```{r acc_vis}

# Build core figures here
#core figures for 50_O and 50_M

bp_labels <- max_accmag_bp_nadir %>%
  group_by(sensor) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

max_accmag_boxplot_nadir_bp <- ggplot(max_accmag_bp_nadir, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = sensor),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = sensor),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = sensor,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("backpack" = "red1", "BDS" = "pink4")) +
  scale_fill_manual(values = c("backpack" = "red1", "BDS" = "pink4")) +
  geom_text(
    data = bp_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~sensor,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(sensor = c(
      "backpack" = "backpack sensors",
      "BDS" = "BDS sensors"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")

#max_accmag_boxplot_nadir_bp


exceedance_plot_nadir_bp <- ggplot(exceed_bp_nadir, aes(x = exceed_pct, y = accmag, group = sensor)) +
  geom_line(aes(color = sensor, linetype = sensor), linewidth = 1) +
  scale_color_manual(values = c("backpack" = "red1", "BDS" = "pink4")) +
  scale_linetype_manual(values = c("backpack" = "solid", "BDS" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Sensor"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")

#exceedance_plot_nadir_bp

combined_plot_nadir_bp <-plot_grid(max_accmag_boxplot_nadir_bp, exceedance_plot_nadir_bp,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")
#combined_plot_nadir_bp

max_accmag_boxplot_post_bp <- ggplot(max_accmag_bp_post, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = sensor),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = sensor),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = sensor,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("backpack" = "steelblue4",
                                "BDS" = "steelblue2")) +
  scale_fill_manual(values = c("backpack" = "steelblue4", 
                               "BDS" = "steelblue2")) +
  geom_text(
    data = bp_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~sensor,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(csensor = c(
      "backpack" = "backpack sensors",
      "BDS" = "BDS sensors"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
  #coord_cartesian(clip = "off")

#max_accmag_boxplot_post

exceedance_plot_post_bp <- ggplot(exceed_bp_post, aes(x = exceed_pct, y = accmag, group = sensor)) +
  geom_line(aes(color = sensor, linetype = sensor), linewidth = 1) +
  scale_color_manual(values = c("backpack" = "steelblue4", "BDS" = "steelblue2")) +
  scale_linetype_manual(values = c("backpack" = "solid", "BDS" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Sensor"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
coord_cartesian(clip = "off")

#exceedance_plot_post

combined_plot_post_nadir_bp <-plot_grid(max_accmag_boxplot_post_bp,
                                     exceedance_plot_post_bp,
                          ncol = 2, nrow = 1, rel_widths = c(0.6,0.4),
                          align = "h",axis = "tb")
#combined_plot_post_nadir_bp



combined_acc_plot_bp <-plot_grid(combined_plot_nadir_bp, 
                              combined_plot_post_nadir_bp,
                              ncol = 1, nrow = 2,
                              rel_heights =  c(0.465,0.535), align = "h",axis = "tb")
combined_acc_plot_bp


#ggsave("combined_acc_plot_bp.svg",
     #  combined_acc_plot_bp,
      # units = "cm",
      # width = 16,
      # height = 16)


```



##7.13 backpack acceleration data analyis

```{r acc_bp_comparison_BDS_analysis}

#consider the difference in sample size
#backpack n =14
#BDS n=115

ad.test(
  max_accmag_bp_nadir$max_accmag[max_accmag_bp_nadir$
                                     sensor == "backpack"],
  max_accmag_bp_nadir$max_accmag[max_accmag_bp_nadir$
                                     sensor == "BDS"]
  )

wilcox.test(max_accmag ~ sensor,
            max_accmag_bp_nadir)


ad.test(
  exceed_bp_nadir$accmag[exceed_bp_nadir$
                             sensor == "backpack"],
  exceed_bp_nadir$accmag[exceed_bp_nadir$
                             sensor == "BDS"]
  )

ad.test(
  max_accmag_bp_post$max_accmag[max_accmag_bp_post$
                                   sensor == "backpack"],
  max_accmag_bp_post$max_accmag[max_accmag_bp_post$
                                    sensor == "BDS"]
  )

wilcox.test(max_accmag ~ sensor,
            max_accmag_bp_post)

ad.test(
  exceed_bp_post$accmag[exceed_bp_post$
                           sensor == "backpack"],
  exceed_bp_post$accmag[exceed_bp_post$
                          sensor == "BDS"]
  )


```