---
title: "VLH_turbine_090226"
output: html_document
date: "2026-02-09"
editor_options: 
  markdown: 
    wrap: 72
---

This markdown file is formatted to be executed with a single run by importing pre-processed BDS sensor data and performing downstream analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  fig.align = "center"
)
knitr::opts_knit$set(root.dir = getwd())
```

# 1. Required libraries and helpers

```{r libraries, echo=FALSE, message = FALSE}

library(tidyverse) # mainly for dplyr and ggplot
library(plotly) # interactive plot
library(ggdist) # kernel density plot
library(cowplot) #arrange multipanel plots
library(kSamples) # Anderson Darling test
library(binom) #95% CI Wilson
library(dlookr)

```

## 1.1 Import sensor data

Import and clean pre-processed sensor data. Import data already has 'bed sensors' removed. Some final smaller changes made to data. See comments where necessary. 

```{r data_import, echo=FALSE, message = FALSE}

denver_files <- list.files(
  "./input",
  pattern = "^denver_.*\\.csv$",
  full.names = TRUE
)

denver_data <- denver_files %>%
  map_dfr(read_csv)


# Setup data structure with expanded treatments
denver_data <- denver_data %>%
  mutate(
    # Extract RPM as numeric for dose-response modeling
    rpm = as.numeric(str_extract(treatment, "\\d+")),
    
    # Extract configuration type
    configuration = factor(
      str_extract(treatment, "(original|modified)"),
      levels = c("original", "modified"),
      labels = c("Original", "Modified")
    ),
    
    treatment = factor(
      treatment,
      levels = c("30rpm_original", "40rpm_original", "50rpm_original", "60rpm_original",
                 "40rpm_modified", "50rpm_modified"),
      labels = c("30 RPM Original", "40 RPM Original", "50 RPM Original", "60 RPM Original",
                 "40 RPM Modified", "50 RPM Modified")
    ),
    roi = factor(
      roi,
      levels = c("unlabeled", "nadir", "1s_post"),
      labels = c("Unlabelled", "Nadir", "Post-nadir")
    )
  )

# modify 5 sensors... 
denver_data <- denver_data %>%
  filter(
    # Remove C550327103129 entirely
    long_id != "C550327103129",
    # Trim specific sensors
    !(long_id == "U120327084613" & time > 113.51),
    !(long_id == "U100327103204" & time > 422),
    !(long_id == "U160510111356" & time > 85),
    !(long_id == "U190327103259" & time > 290)
  )

```


Input helper functions 

```{r helper_functions, echo=FALSE, message = FALSE}

# setup custom theme
theme_custom <- function(base_size=10){ 
  #base font size set here (10 pt)
  theme_grey() %+replace%
    theme(
      axis.text = element_text(colour="black"),
      axis.title = element_text(colour="black"),
      axis.ticks = element_line(colour="black"),
      panel.border = element_rect(colour = "black", fill=NA),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      strip.background = element_rect(colour = "black", fill = NA),
      panel.spacing.x = unit(12, "pt")
      #default family = sans (Arial)
    ) 
}

```


# 2. Exploratory data analysis

All EDA should be provided here. This includes measuring and validating data dimensions e.g., number sensors for each operational scenario, cleaning data including missing and outlier data and project specific conditions, setting definitions for what data is used for downstream analysis, early summarisations and visualizations which lead into later formal analysis


E.g., 

```{r data_structure, echo=FALSE, message = FALSE}

#Check unique sensor ID for each treatment
denver_data %>%
  group_by(treatment) %>%
  summarise(
    unique_long_ids = n_distinct(long_id),
    .groups = "drop"
  )

str(denver_data)

colnames(denver_data)

dlookr::plot_outlier(denver_data)

```

## 2.1 Visualise main relationships


By plotting the main visual relationships we're able to quickly identify relationships in the data and begin addressing study objectives and hypothesis.

Think of your study objectives.... what do those relationships look like?

E.g.,

```{r}

plot_data_post <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time),
         mean_strain = round(mean_strain)
         ) %>%
  ungroup()

# Get max acceleration points for each sensor
max_points <- plot_data_post %>%
  group_by(long_id, treatment) %>%
  slice_max(order_by = abs(accmag), n = 1, with_ties = FALSE) %>%
  ungroup()

# accmag_time_plot
accmag_post_nadir <- ggplot(plot_data_post, aes(x = time_post, y = abs(accmag), group = long_id, color = treatment)) +
  geom_line(alpha = 0.5, color = "black") +
  geom_point(data = max_points, alpha = 0.7, size = 2, color = "red2") +
  facet_wrap(~ treatment, nrow = 1, labeller = labeller(treatment = c(
    "Original" = "Original Configuration",
    "Modified" = "Modified Configuration"
  ))) +
  labs(
    x = " Time (s) (ROI 3 Post-nadir)", 
    y = expression("Acceleration Magnitude (m s" ^-2*")")) +
  scale_x_continuous(breaks = seq(0, 1, 0.1), expand = c(0,0))+
  scale_y_continuous(breaks = seq(0, 250, 25), expand= c(0.01,0.01))+
  theme_custom() +
  theme(axis.text.x = element_text(angle=90),
        legend.position = "none")

accmag_post_nadir

```

### 2.1.1 Example passage plot

```{r}

#### Create example passage plot 

C39 <- denver_data %>%
  filter (long_id == "C390327085054")

# # Plot plotly to identify trim times etc
# plot_ly(C39, x = ~time) %>%
#   add_lines(y = ~pres, name = "Pres", line = list(color = "black"), yaxis = "y") %>%
#   add_lines(y = ~accmag, name = "Acc", line = list(color = "red"), yaxis = "y2") %>%
#   layout(
#     xaxis = list(title = "Time (s)"),
#     yaxis = list(title = "Pressure (kPa)", titlefont = list(color = "black")),
#     yaxis2 = list(title = "Acceleration Magnitude (g)", titlefont = list(color = "red"),
#                   overlaying = "y", side = "right")
#   )

# decided to Trim data again to 106-121
# ~ 3s plot
C39 <- C39 %>%
  filter(time >= 65.598, time <= 69)

# Identify ROI boundaries 
roi_changes <- C39 %>%
  mutate(roi_change = roi != lag(roi, default = first(roi))) %>%
  filter(roi_change) %>%
  pull(time)

#Identify accmag and pressure min max for secondary axis scaling
pres_min <- 90
pres_max <- 120
accmag_min <- min(C39$accmag)  # approximately -5.809
accmag_max <- 230

# Calculate scaling to map accmag range onto pressure range
scale_factor <- (pres_max - pres_min) / (accmag_max - accmag_min)
offset <- pres_min - accmag_min * scale_factor


C39_plot <- ggplot(C39, aes(x = time)) +
  geom_rect(data = C39 %>%
              summarise(xmin = min(time), xmax = max(time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey90", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(data = C39 %>% filter(roi == "Nadir") %>% 
              summarise(xmin = min(time), xmax = max(time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#F4A6A3", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(data = C39 %>% filter(roi == "Post-nadir") %>% 
              summarise(xmin = min(time), xmax = max(time)),
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "#A8DADC", alpha = 0.1, inherit.aes = FALSE) +
  geom_line(aes(y = pres), color = "black", size = 0.5) +
  geom_line(aes(y = accmag * scale_factor + offset), color = "red2", size = 0.5) +
  geom_vline(xintercept = roi_changes, color = "blue4", linetype = "dotted") +
  scale_x_continuous(breaks = seq(65, 69, by = 0.25), expand = c(0,0)) +
  scale_y_continuous(
    name = "Pressure (kPa)",
    breaks = seq(90, 120, by = 2),
    limits = c(90, 120),
    sec.axis = sec_axis(
      ~ (. - offset) / scale_factor, 
      name = expression("Acceleration Magnitude (m s" ^-2*")"),
      breaks = seq(0, 230, by = 20)
    )
  ) +
  labs(x = "Time (s)") +
  theme_custom() +
  theme(
    axis.title.y.right = element_text(color = "red2"),
    axis.text.y.right = element_text(color = "red2"),
    axis.text.x = element_text(angle = 90)
  ) +
  coord_cartesian(clip = "off")

C39_plot

# Export svg for editing in Inkscape. Ensure units are cm and size appropriate for page space. No DPI needed (vector)
#ggsave("C39_plot.svg", C39_plot, units = "cm", width = 16, height = 10)

# Export png. Ensure units are cm and size appropriate for page space. Set DPI at 300
#ggsave("C39_plot.png", C39_plot, dpi = 300, units = "cm", width = 16, height = 10)



```




# 3. Pressure indicators

Data preparation 

```{r pres_prep}

# Filter and shape data accordingly for pressure analysis


```

## 3.1 Pressure summary


```{r pres_sum}

# Summarise and tabulate pressure variables across treatments etc


```


## 3.2 Pressure vis


```{r pres_vis}

# Build core figures here


```



## 3.3 Pressure analysis

```{r pres_analysis}

# Build core figures here


```



# 4. Acceleration indicators

Data preparation 

```{r acc_prep}

# Filter and shape data accordingly for acceleration analysis

max_accmag_data_nadir <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Nadir") %>%
  group_by(long_id, ) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")

plot_data_nadir <- denver_data %>% 
  filter(rpm == 50) %>%
  filter(roi == "Nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time)) %>%
  ungroup()

exceed_data_nadir <- plot_data_nadir %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()


# Get max accmag for each sensor in Post-nadir ROI
max_accmag_data_post <- denver_data %>% 
  filter(rpm == 50) %>%
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  summarise(max_accmag = max(abs(accmag)), .groups = "drop")


plot_data_post <- denver_data %>%
  filter(rpm == 50) %>% 
  filter(roi == "Post-nadir") %>%
  group_by(long_id, configuration) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(time_post = time - min(time),
         mean_strain = round(mean_strain)
         ) %>%
  ungroup()

exceed_data_post <- plot_data_post %>%
  mutate(accmag = abs(accmag)) %>%
  filter(accmag > 10) %>%
  group_by(configuration) %>%
  arrange(accmag, .by_group = TRUE) %>%
  mutate(
    p_rank = percent_rank(accmag),
    exceed_pct = (1 - p_rank) * 100
  ) %>%
  ungroup()

```

## 4.1 Acceleration summary


```{r acc_sum}

# Summarise and tabulate acceleration variables across treatments etc

#e.g.,

#central tendencies
max_accmag_data_post %>%
  group_by(configuration) %>%
  summarise(
    n = n(),
    median = median(max_accmag),
    Q1 = quantile(max_accmag, 0.25),
    Q3 = quantile(max_accmag, 0.75),
    IQR = IQR(max_accmag),
    mean = mean(max_accmag),
    sd = sd(max_accmag),
    .groups = "drop"
  )



```


## 4.2 Acceleration vis


```{r acc_vis}

# Build core figures here

n_labels <- max_accmag_data_nadir %>%
  group_by(configuration) %>%
  summarise(n = n_distinct(long_id), .groups = "drop")

max_accmag_boxplot_nadir <- ggplot(max_accmag_data_nadir, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_fill_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
  )
#coord_cartesian(clip = "off")

#max_accmag_boxplot_nadir


exceedance_plot_nadir <- ggplot(exceed_data_nadir, aes(x = exceed_pct, y = accmag, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "red1", "Modified" = "pink4")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_blank(), axis.title.x = element_blank())+
  coord_cartesian(clip = "off")

#exceedance_plot_nadir

combined_plot_nadir <-plot_grid(max_accmag_boxplot_nadir, exceedance_plot_nadir,
                                ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")
#combined_plot_nadir

max_accmag_boxplot_post <- ggplot(max_accmag_data_post, aes(y = max_accmag)) +
  stat_halfeye(
    aes(fill = configuration),
    side = "right",
    adjust = 0.5,
    slab_alpha = 0.2,
    slab_linewidth = 0.8,
    .width = 0
  ) +
  geom_boxplot(
    aes(color = configuration),
    width = 0.2,
    outlier.shape = 4, 
    orientation = "horizontal",
    staplewidth = 0.5,
    linewidth = 0.8,
    outlier.size = 2.5,
    outlier.stroke = 0.8
  ) +
  geom_jitter(
    aes(color = configuration,x = 0),
    width = 0.15,
    height = 0,
    alpha = 0.4,
    size = 1
  )+
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_fill_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  geom_text(
    data = n_labels,
    aes(label = paste0("sensor (n) = ", n), x = 0.95, y = 200)
  )+
  coord_flip() +
  facet_wrap(
    ~configuration,
    ncol = 1,
    strip.position = "top",
    labeller = labeller(configuration = c(
      "Original" = "Original Configuration",
      "Modified" = "Modified Configuration"
    ))
  ) +
  labs(
    x = "Kernel Density",
    y = expression("Max Acceleration Magnitude (m s"^-2*")")
  ) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0,230)) +
  theme_custom() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
  )
  #coord_cartesian(clip = "off")

#max_accmag_boxplot_post

exceedance_plot_post <- ggplot(exceed_data_post, aes(x = exceed_pct, y = accmag, group = configuration)) +
  geom_line(aes(color = configuration, linetype = configuration), linewidth = 1) +
  scale_color_manual(values = c("Original" = "steelblue4", "Modified" = "steelblue2")) +
  scale_linetype_manual(values = c("Original" = "solid", "Modified" = "solid")) +
  scale_y_continuous(breaks = seq(0, 230, 25), limits = c(0, 230)) +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100), expand = c(0.01, 0)) +
  labs(
    x = "Percentage Exceedance (%)",
    y = expression("Acceleration Magnitude (m s"^-2*")"),
    color = "Treatment"
  ) +
  theme_custom() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
coord_cartesian(clip = "off")

#exceedance_plot_post

combined_plot_post_nadir <-plot_grid(max_accmag_boxplot_post, exceedance_plot_post,
                          ncol = 2, nrow = 1, rel_widths = c(0.6,0.4), align = "h",axis = "tb")
combined_plot_post_nadir

#ggsave("combined_plot_post_nadir.svg", combined_plot_post_nadir, units = "cm", width = 16, height = 8)





combined_acc_plot <-plot_grid(combined_plot_nadir, combined_plot_post_nadir,
                                     ncol = 1, nrow = 2, rel_heights =  c(0.465,0.535), align = "h",axis = "tb")
combined_acc_plot


ggsave("combined_acc_plot.svg", combined_acc_plot, units = "cm", width = 16, height = 16)


```



## 4.3 Acceleration analysis

```{r acc_analysis}

# Core analysis here

ad.test(
  max_accmag_data_nadir$max_accmag[max_accmag_data_nadir$treatment == "Original"],
  max_accmag_data_nadir$max_accmag[max_accmag_data_nadir$treatment == "Modified"]
)

wilcox.test(max_accmag ~ configuration, max_accmag_data_nadir)

ad.test(
  exceed_data_nadir$accmag[exceed_data_nadir$treatment == "Original"],
  exceed_data_nadir$accmag[exceed_data_nadir$treatment == "Modified"]
)



ad.test(
  max_accmag_data_post$max_accmag[max_accmag_data_post$treatment == "Original"],
  max_accmag_data_post$max_accmag[max_accmag_data_post$treatment == "Modified"]
)

wilcox.test(max_accmag ~ configuration, max_accmag_data_post)

ad.test(
  exceed_data_post$accmag[exceed_data_post$treatment == "Original"],
  exceed_data_post$accmag[exceed_data_post$treatment == "Modified"]
)


```


# 5. Strain indicators

Data preparation 

```{r strain_prep}

# Filter and shape data accordingly for pressure analysis


```

## 5.1 Strain summary


```{r strain_sum}

# Summarise and tabulate pressure variables across treatments etc


```


## 5.2 Strain vis


```{r strain_vis}

# Build core figures here


```



## 5.3 Strain analysis

```{r strain_analysis}

# Core analysis here


```

# 6. Fish survival estimates


```{r}

survival_data <- tibble(
  species = c(
    "Roach (Rutilus rutilus)",
    "European eel (Anguilla anguilla) - adult",
    "European eel (Anguilla anguilla) - juvenile",
    "Bream (Abramis brama)",
    "Brown trout (Salmo trutta)"
  ),
  total_recap = c(239, 46, 219, 208, 211),
  total_surv  = c(237, 46, 219, 207, 211),
  group       = NA,
  survival    = NA,
  lower_ci    = NA,
  upper_ci    = NA
)

survival__ci <- survival_data %>%
  mutate(
    # Calculate survival proportion
    survival = total_surv / total_recap,
    # Calculate Wilson CIs
    ci_result = map2(total_surv, total_recap, 
                     ~binom.confint(.x, .y, methods = "wilson"))
  ) %>%
  unnest(ci_result) %>%
  select(species, group, total_recap, total_surv, 
         survival = mean, lower_ci = lower, upper_ci = upper)

# Display results
print(survival__ci)

```

```{r}

impact_data <- survival__ci %>%
  mutate(species = factor(species, levels = rev(unique(species))))

p <- ggplot(impact_data, aes(x = survival * 100, y = species)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, 
           fill = "#ffcccc", alpha = 0.3) +
  annotate("rect", xmin = 96, xmax = 100, ymin = -Inf, ymax = Inf, 
           fill = "#ccffcc", alpha = 0.5) +
  geom_vline(xintercept = c(96, 98, 100), linetype = "dashed", 
             color = "grey60", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = lower_ci * 100, xmax = upper_ci * 100), 
                 height = 0.2, linewidth = 0.5) +
  geom_point(size = 2, shape = 16) +
  geom_text(aes(label = paste0("italic(n) == ", total_recap)), 
            x = 90.5, hjust = 0, size = 3.5, parse = TRUE) +
  # Top left label
  annotate("text", x = 90, y = Inf, 
           label = "CI = 95% (Wilson Binomial)\nMargin = 2%",
           hjust = 0, vjust = 1, size = 3.5, lineheight = 0.9) +
  # Labels and theme
  scale_x_continuous(limits = c(90, 100), 
                     breaks = seq(90, 100, 1)) +
  labs(x = "Survival Probability (%)",
       y = NULL) +
  theme_custom() +
  coord_cartesian(clip = "off")

print(p)

ggsave("denver_ci.svg", 
       plot = p, 
       width = 18, 
       height = 10, 
       units = "cm",
       bg = "white")

```



